A Plot is Worth a Thousand Tests: Assessing Residual Diagnostics with the Lineup Protocol

### 2023-05-04 paper.tex
final changes

\begin{abstract}
Regression experts consistently recommend plotting residuals for model
diagnosis, despite the availability of many numerical hypothesis test
procedures designed to use residuals to assess problems with a model
fit. Here we provide evidence for why this is good advice using data
from a visual inference experiment. We show how conventional tests are
too sensitive, which means that too often the conclusion would be that
the model fit is inadequate. The experiment uses the lineup protocol
which puts a residual plot in the context of null plots. This helps
generate reliable and consistent reading of residual plots for better
model diagnosis. It can also help in an obverse situation where a
conventional test would fail to detect a problem with a model due to
contaminated data. The lineup protocol also serves as an omnibus test,
capable of detecting a range of departures from good residuals
simultaneously.
\end{abstract}

@@

We can assess residuals in multiple ways. To examine the their
univariate distribution, residuals may be plotted as a histogram or
normal probability plot. Using the classical normal linear regression
model as an example, if the distribution is symmetric and unimodal, we
would consider it to be well-behaved. However, if the distribution is
skewed, bimodal, multimodal, or contains outliers, there would be cause
for concern. We can also inspect the distribution by conducting a
goodness-of-fit test, such as the Shapiro-Wilk normality test
\citep{shapiro1965analysis}.

In addition, scatterplots of residuals plotted against the fitted values
and each of the explanatory variables, can be used to scrutinize their
relationships. If there are any visually discoverable patterns, the
model is potentially inadequate or incorrectly specified. In general,
one looks for noticeable departures from the model such as non-linear
pattern or heteroskedasticity. A non-linear pattern would suggest that
the model needs to have some additional non-linear terms.
Heteroskedasticity suggests that the error is dependent on the
predictors, and hence violates the independence assumption. However,
correctly judging whether NO pattern exists in a residual plot is a
difficult task for humans. Humans will almost always will see a pattern
(see \citet{kahneman2011thinking}), so the question that really needs
answering is whether any pattern perceived is consistent with
randomness, purely sampling variability, or noise. It is especially
difficult to teach this to new analysts and students
\citep{loy2021bringing}. To address this, statistical tests are
available to check for non-linear patterns
\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
\citep[e.g.][]{breusch1979simple}.

@@

Model misspecification occurs if functions of predictors that needed to
accurately describe the relationship between response and predictors are
omitted. Any non-linear pattern visible in the residual plot could be
indicative of this problem. An example residual plot containing visual
pattern of non-linearity is shown in Figure
\ref{fig:residual-plot-common-departures}B. One can clearly observe the
``S-shape'' from the residual plot, which corresponds to cubic term that
should have been included in the model.

@@

Many different hypothesis tests are available to detect specific model
defects. For example, the presence of heteroskedasticity can usually be
tested by applying the White test \citep{white1980heteroskedasticity} or
the Breusch-Pagan (BP) test \citep{breusch1979simple}, which are both
derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
principle that relies on the asymptotic properties of the null
distribution. To test specific forms of non-linearity, one may apply the
F-test as a model structural test to examine the significance of
specific polynomial and non-linear forms of the predictors, or the
significance of proxy variables as in the Ramsey Regression Equation
Specification Error Test (RESET) \citep{ramsey1969tests}. The
Shapiro-Wilk (SW) normality test \citep{shapiro1965analysis} is the most
widely used test of non-normality included by many of the statistical
software programs. The Jarque-Bera test \citep{jarque1980efficient} is
also used to directly check whether the sample skewness and kurtosis
match a normal distribution.

Table \ref{tab:example-residual-plot-table} displays the \(p\)-values
from the RESET, BP and SW tests applied to the residual plots in Figure
\ref{fig:residual-plot-common-departures}. The RESET test and BP test
were computed using the \texttt{resettest} and \texttt{bptest} functions
from the R package \texttt{lmtest}, respectively. The SW test was
computed using the \texttt{shapiro.test} from the core R package
\texttt{stats}. \footnote{Although we didn't use it, it's useful to know
  that the R package \texttt{skedastic} \citep{skedastic} also contains
  a large collection of functions to test for heteroskedasticity.}
Although, the RESET test is exact, it requires the selection of a power
parameter. \citet{ramsey1969tests} recommends a power of four, which we
adopted in our analysis. The BP and SW tests are approximate.

We would expect the RESET test for non-linearity to reject residual plot
B, the BP test for heteroskedasticity to reject the residual plot C, and
the SW test for non-normality to reject residual plot D, which they all
do and all tests also correctly fail to reject residual plot A.
Interestingly, the BP and SW tests also reject the residual plots
exhibiting structure that they weren't designed for.
\citet{cook1982residuals} explain that most residual-based tests for
particular types of departure from model assumptions are also sensitive
to other types of departures. This could be considered a Type III error
\citep{kimball1957errors}, where the null hypothesis of good residuals
is correctly rejected but for the wrong reason. Also, some types of
departure can have elements of other types of departure, for example,
non-linearity could be viewed as heteroskedasticity. Additionally, other
data problems such as outliers can trigger rejection (or not) of the
null hypothesis \citep{cook1999applied}.

@@

The examination of data plots to infer signal or patterns (or lack
thereof) is fraught with variation in the human ability to interpret and
decode the information embedded in a graph
\citep{cleveland1984graphical}. Human examination of diagnostic plots
can feel subjective. Ideally we would expect visual evaluation to be
done objectively with small amounts of human-to-human variation in
reading data plots, but we cannot assume this is the case.

In practice, over-interpretation of a single plot is common. For
instance, \citet{roy2015using} described a published example where
authors over-interpreted separation between gene groups from a
two-dimensional projection of a linear discriminant analysis.
\citet{roy2015using} showed that there were no differences in the
expression levels between the gene groups. One solution to
over-interpretation is to examine the plot in the context of natural
sampling variability assumed by the model, called the lineup protocol,
as proposed in \citet{buja2009statistical}.
\citet{majumder2013validation} showed that the lineup protocol is
analogous to the null hypothesis significance testing framework. The
protocol consists of \(m\) randomly placed plots, where one plot is the
data plot, and the remaining \(m - 1\) plots, referred to as the
\emph{null plots}, are constructed using the same graphical procedure as
the data plot but the data is replaced with null data that is generated
in a manner consistent with the null hypothesis, \(H_0\). Then, an
observer who has not seen the data plot is asked to point out the most
different plot from the lineup. Under \(H_0\), it is expected that the
data plot would have no distinguishable difference from the null plots,
and the probability that the observer correctly picks the data plot is
\(1/m\). If one rejects \(H_0\) as the observer correctly picks the data
plot, then the Type I error of this test is \(1/m\). This protocol
requires a priori specification of \(H_0\) (or at least a null data
generating mechanism), much like the requirement of knowing the sampling
distribution of the test statistic in null hypothesis significance
testing framework.

@@

Data used in the \(m - 1\) null plots needs to be simulated. In
regression diagnostics, sampling data consistent with \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja2009statistical} suggested, \(H_0\) is usually a composite
hypothesis controlled by nuisance parameters. Since regression models
can have various forms, there is no general solution to this problem,
but it sometimes can be reduced to a so called ``reference
distribution'' by applying one of the three methods: (i) sampling from a
conditional distribution given a minimal sufficient statistic under
\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
estimated under \(H_0\), and (iii) Bayesian posterior predictive
sampling. The conditional distribution given a minimal sufficient
statistic is the best justified reference distribution among the three
\citep{buja2009statistical}. Essentially, null residuals can be
simulated by regressing \(N\) i.i.d standard normal random draws on the
predictors, then re-scaling it by the ratio of residual sum of square in
two regressions.

The effectiveness of lineup protocol for regression analysis has been
validated by \citet{majumder2013validation} under relatively simple
settings with up to two predictors. Their results suggest that visual
tests are capable of testing the significance of a single predictor with
a similar power to a t-test, though they express that in general it is
unnecessary to use visual inference if there exists a conventional test,
and they do not expect the visual test to perform equally well as the
conventional test. In their third experiment, where there is not a
conventional test, visual test outperforms the conventional test by a
large margin. This is encouraging, as it supports the use of visual
inference in situations where there are no existing statistical testing
procedures. Visual inference has also been integrated into diagnostics
for hierarchical linear models where the lineup protocol is used to
judge the assumptions of linearity, normality and constant error
variance for both the level-1 and level-2 residuals
\citep[\citet{loy2014hlmdiag} and
\citet{loy2015you}]{loy2013diagnostic}.

\hypertarget{calculation-of-statistical-significance-and-test-power}{%
\section{Calculation of statistical significance and test
power}\label{calculation-of-statistical-significance-and-test-power}}

@@

As pointed out by \citet{vanderplas2021statistical}, this basic binomial
model is deficient. It does not take into account the possible
dependencies in the visual test due to repeated evaluations of the same
lineup, or account for when participants are offered the option to
select one or more ``most different'' plots, or none, from a lineup.
They suggest three common lineup scenarios: (1) \(K\) different lineups
are shown to \(K\) participants, (2) \(K\) lineups with different null
plots but the same data plot are shown to \(K\) participants, and (3)
the same lineup is shown to \(K\) participants. Scenario 3 is the most
feasible to apply, but has the most dependencies to accommodate for the
\(p\)-value calculation. For Scenario 3, VanderPlas et al propose
modelling the probability of plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
selected in \(K\) evaluations is denoted as \(c_i\). In case participant
\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
instead of one, where \(s_j\) is the number of plots participant \(j\)
selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
only interested in the selections of the data plot \(i\), the marginal
model can be simplified to a beta-binomial model and thus the visual
\(p\)-value is given as

@@

The parameter \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial}
and \ref{eq:pmf-beta-binomial} is usually unknown and will need to be
estimated from the survey data. An interpretation of \(\alpha\) is that
when it is low only a few plots are attractive to the observers and tend
to be selected, and when high, most plots are equally likely to be
chosen. VanderPlas et al define \(c\)-interesting plot to be if \(c\) or
more participants select the plot as the most different. The expected
number of plots selected at least \(c\) times, \(E[Z_c]\), is then
calculated as

\begin{equation} \label{eq:c-interesting-expectation}
E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}

@@

Visual patterns of non-linearity are simulated using four different
orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). The
values of \(j\) were chosen so that distinct shapes of non-linearity
were included in the residual plot. These include ``U'', ``S'', ``M''
and ``triple-U'' shape as shown in Figure
\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
result in a curve with more turning points. It is expected that the
``U'' shape will be the easiest detect, as the shape gets more complex
it will be harder to perceive in a scatterplot, particularly when there
is noise. Figure \ref{fig:different-sigma} shows the ``U'' shape for
different amounts of noise (\(\sigma\)).

@@

Since \(supp(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\) and
\(1\) can generate ``left-triangle'', ``butterfly'' and
``right-triangle'' shapes as displayed in Figure
\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
magnitude of residuals across different values of \(a\). Figure
\ref{fig:different-b} shows the butterfly shape as the ratio parameter
(\(b\)) is changed.

@@

\begin{itemize}
\tightlist
\item
  uniform, \(U(-1, 1)\),
\item
  normal, \(N(0, 0.3^2)\),
\item
  skewed, \(lognormal(0, 0.6^2)/3\), and
\item
  discrete, \(U\{1, 5\}\).
\end{itemize}

@@

\begin{equation}
E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(diag(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)
\label{eq:nonlin-effect-size}
\end{equation}

\noindent where \(diag(.)\) is the diagonal matrix constructed from the
diagonal elements of a matrix,
\(\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}\) is the residual
operator,
\(\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
is the hat matrix,
\(\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of residuals where \(\boldsymbol{Z}\) contain any
higher order terms of \(\boldsymbol{X}\) left out of the regression
equation, \(\boldsymbol{\beta}_z\) contains the corresponding
coefficients, and \(\sigma^2\boldsymbol{I}\) is the assumed covariance
matrix of the error term when \(H_0\) is true.

@@

Data collection used a total of 1152 lineups, and resulted in a total of
7974 evaluations from 443 participants. Roughly half corresponded to the
two models, non-linearity and heteroskedasticiy, and the three
collection periods had similar numbers of evaluations. Each participant
received two of the 24 attention check lineups which were used to filter
results of participants who were clearly not making an honest effort
(only 11 of 454). To estimate \(\alpha\) for calculating statistical
significance (see Section \ref{sig}) there were 720 evaluations of 36
null lineups. Neither the attention checks nor null lineups were used in
the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
made available in the R package, \texttt{visage}.\footnote{The R package
  \texttt{visage} is currently available on GitHub
  (\url{https://github.com/TengMCing/visage}).}

@@

In regression analysis, predictions are conditional on the observed
values of the predictors, that is, the conditional mean of the dependent
variable \(Y\) given the value of the independent variable \(X\),
\(E(Y|X)\). This is an often forgotten element of regression analysis
but it is important. Where \(X\) is observed, the distribution of the
\(X\) values in the sample, or consequently \(\hat{Y}\), may affect the
ability to read any patterns in the residual plots. The effect of fitted
value distribution on test performance is assess using four different
distributions of fitted values: uniform, normal, discrete and lognormal
(skewed). We expect that if \(\hat{Y}\) has a uniform distribution, it
is easier to read the relationship with the residuals.

@@

>>>>

\begin{abstract}
Regression experts consistently recommend plotting residuals for model
diagnosis, despite the availability of many numerical hypothesis test
procedures designed to use residuals to assess problems with a model
fit. Here we provide evidence for why this is good advice using data
from a visual inference experiment. We show how conventional tests are
too sensitive, which means that too often the conclusion would be that
the model fit is inadequate. The experiment uses the lineup protocol
which puts a residual plot in the context of null plots. This helps
generate reliable and consistent reading of residual plots for better
model diagnosis. It can also help in an obverse situation where a
conventional test would fail to detect a problem with a model due to
contaminated data. The lineup protocol also detects a range of
departures from good residuals simultaneously.
\end{abstract}

@@

We can assess residuals in multiple ways. To examine the univariate
distribution, residuals may be plotted as a histogram or normal
probability plot. Using the classical normal linear regression model as
an example, if the distribution is symmetric and unimodal, we would
consider it to be well-behaved. However, if the distribution is skewed,
bimodal, multimodal, or contains outliers, there would be cause for
concern. We can also inspect the distribution by conducting a
goodness-of-fit test, such as the Shapiro-Wilk normality test
\citep{shapiro1965analysis}.

In addition, scatterplots of residuals plotted against the fitted values
and each of the explanatory variables, can be used to scrutinize their
relationships. If there are any visually discoverable patterns, the
model is potentially inadequate or incorrectly specified. In general,
one looks for noticeable departures from the model such as non-linear
pattern or heteroskedasticity. A non-linear pattern would suggest that
the model needs to have some additional non-linear terms.
Heteroskedasticity suggests that the error is dependent on the
predictors, and hence violates the independence assumption. However,
correctly judging whether NO pattern exists in a residual plot is a
difficult task for humans. Humans will almost always see a pattern (see
\citet{kahneman2011thinking}), so the question that really needs
answering is whether any pattern perceived is consistent with
randomness, purely sampling variability, or noise. It is especially
difficult to teach this to new analysts and students
\citep{loy2021bringing}. To address this, statistical tests are
available to check for non-linear patterns
\citep[e.g.][]{ramsey1969tests}, and heteroskedasticity
\citep[e.g.][]{breusch1979simple}.

@@

Model misspecification occurs if functions of predictors that needed to
accurately describe the relationship with the response are incorrectly
specified. This includes instances where a higher-order polynomial term
of a predictor is wrongfully omitted. Any non-linear pattern visible in
the residual plot could be indicative of this problem. An example
residual plot containing visual pattern of non-linearity is shown in
Figure \ref{fig:residual-plot-common-departures}B. One can clearly
observe the ``S-shape'' from the residual plot, which corresponds to
cubic term that should have been included in the model.

@@

Many different hypothesis tests are available to detect specific model
defects. For example, the presence of heteroskedasticity can usually be
tested by applying the White test \citep{white1980heteroskedasticity} or
the Breusch-Pagan (BP) test \citep{breusch1979simple}, which are both
derived from the Lagrange multiplier test \citep{silvey1959lagrangian}
principle that relies on the asymptotic properties of the null
distribution. To test specific forms of non-linearity, one may apply the
F-test as a model structural test to examine the significance of a
specific polynomial and non-linear forms of the predictors, or the
significance of proxy variables as in the Ramsey Regression Equation
Specification Error Test (RESET) \citep{ramsey1969tests}. The
Shapiro-Wilk (SW) normality test \citep{shapiro1965analysis} is the most
widely used test of non-normality included by many of the statistical
software programs. The Jarque-Bera test \citep{jarque1980efficient} is
also used to directly check whether the sample skewness and kurtosis
match a normal distribution.

Table \ref{tab:example-residual-plot-table} displays the \(p\)-values
from the RESET, BP and SW tests applied to the residual plots in Figure
\ref{fig:residual-plot-common-departures}. The RESET test and BP test
were computed using the \texttt{resettest} and \texttt{bptest} functions
from the R package \texttt{lmtest}, respectively. The SW test was
computed using the \texttt{shapiro.test} from the core R package
\texttt{stats}. \footnote{Although we didn't use it, it's useful to know
  that the R package \texttt{skedastic} \citep{skedastic} also contains
  a large collection of functions to test for heteroskedasticity.} The
RESET test requires the selection of a power parameter.
\citet{ramsey1969tests} recommends a power of four, which we adopted in
our analysis.

For residual plots in Figure \ref{fig:residual-plot-common-departures},
we would expect the RESET test for non-linearity to reject residual plot
B, the BP test for heteroskedasticity to reject the residual plot C, and
the SW test for non-normality to reject residual plot D, which they all
do and all tests also correctly fail to reject residual plot A.
Interestingly, the BP and SW tests also reject the residual plots
exhibiting structure that they weren't designed for.
\citet{cook1982residuals} explain that most residual-based tests for a
particular departure from the model assumptions are also sensitive to
other types of departures. This could be considered a Type III error
\citep{kimball1957errors}, where the null hypothesis of good residuals
is correctly rejected but for the wrong reason. Also, some types of
departure can have elements of other types of departure, for example,
non-linearity can appear like heteroskedasticity. Additionally, other
data problems such as outliers can trigger rejection (or not) of the
null hypothesis \citep{cook1999applied}.

@@

The examination of data plots to infer signal or patterns (or lack
thereof) is fraught with variation in the human ability to interpret and
decode the information embedded in a graph
\citep{cleveland1984graphical}. Human examination of diagnostic plots
can feel subjective. Ideally we would expect visual evaluation to be
done objectively with low human-to-human variation in reading data
plots, but we cannot assume this is the case.

In practice, over-interpretation of a single plot is common. For
instance, \citet{roy2015using} described a published example where
authors over-interpreted separation between gene groups from a
two-dimensional projection of a linear discriminant analysis even when
there were no differences in the expression levels between the gene
groups. One solution to over-interpretation is to examine the plot in
the context of natural sampling variability assumed by the model, called
the lineup protocol, as proposed in \citet{buja2009statistical}.
\citet{majumder2013validation} showed that the lineup protocol is
analogous to the null hypothesis significance testing framework. The
protocol consists of \(m\) randomly placed plots, where one plot is the
data plot, and the remaining \(m - 1\) plots, referred to as the
\emph{null plots}, are constructed using the same graphical procedure as
the data plot but the data is replaced with null data that is generated
in a manner consistent with the null hypothesis, \(H_0\). Then, an
observer who has not seen the data plot is asked to point out the most
different plot from the lineup. Under \(H_0\), it is expected that the
data plot would have no distinguishable difference from the null plots,
and the probability that the observer correctly picks the data plot is
\(1/m\). If one rejects \(H_0\) as the observer correctly picks the data
plot, then the Type I error of this test is \(1/m\). This protocol
requires a priori specification of \(H_0\) (or at least a null data
generating mechanism), much like the requirement of knowing the sampling
distribution of the test statistic in null hypothesis significance
testing framework.

@@

Data used in the \(m - 1\) null plots needs to be simulated. In
regression diagnostics, sampling data consistent with \(H_0\) is
equivalent to sampling data from the assumed model. As
\citet{buja2009statistical} suggested, \(H_0\) is usually a composite
hypothesis controlled by nuisance parameters. Since regression models
can have various forms, there is no general solution to this problem,
but it sometimes can be reduced to a so called ``reference
distribution'' by applying one of the three methods: (i) sampling from a
conditional distribution given a minimal sufficient statistic under
\(H_0\), (ii) parametric bootstrap sampling with nuisance parameters
estimated under \(H_0\), and (iii) Bayesian posterior predictive
sampling. The conditional distribution given a minimal sufficient
statistic is the best justified reference distribution among the three
\citep{buja2009statistical}. Under this method, the null residuals can
essentially be simulated by independent drawing from a standard normal
random distribution, then regressing the draws on the predictors, and
then re-scaling it by the ratio of the residual sum of square in two
regressions.

The effectiveness of lineup protocol for regression analysis has been
validated by \citet{majumder2013validation} under relatively simple
settings with up to two predictors. Their results suggest that visual
tests are capable of testing the significance of a single predictor with
a similar power to a t-test, though they express that in general it is
unnecessary to use visual inference if there exists a corresponding
conventional test, and they do not expect the visual test to perform
equally well as the conventional test. In their third experiment, where
the contamination of the data violate the assumptions of the convetional
test, visual test outperforms the conventional test by a large margin.
This is encouraging, as it supports the use of visual inference in
situations where there are no existing numerical testing procedures.
Visual inference has also been integrated into diagnostics for
hierarchical linear models where the lineup protocol is used to judge
the assumptions of linearity, normality and constant error variance for
both the level-1 and level-2 residuals \citep[\citet{loy2014hlmdiag} and
\citet{loy2015you}]{loy2013diagnostic}.

\hypertarget{significance-calculation}{%
\section{Calculation of statistical significance and test
power}\label{significance-calculation}}

@@

As pointed out by \citet{vanderplas2021statistical}, this basic binomial
model is deficient. It does not take into account the possible
dependencies in the visual test due to repeated evaluations of the same
lineup, or account for when participants are offered the option to
select one or more ``most different'' plots, or none, from a lineup.
They suggest three common lineup scenarios: (1) \(K\) different lineups
are shown to \(K\) participants, (2) \(K\) lineups with different null
plots but the same data plot are shown to \(K\) participants, and (3)
the same lineup is shown to \(K\) participants. Scenario 3 is the most
feasible to apply, but has the most dependencies to accommodate for the
\(p\)-value calculation. For Scenario 3, VanderPlas et al propose
modelling the probability of plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim \text{Dirichlet}(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
selected in \(K\) evaluations is denoted as \(c_i\). In case participant
\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
instead of one, where \(s_j\) is the number of plots participant \(j\)
selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
only interested in the selections of the data plot \(i\), the marginal
model can be simplified to a beta-binomial model and thus the visual
\(p\)-value is given as

@@

The parameter \(\alpha\) used in Equation \ref{eq:pvalue-beta-binomial}
and \ref{eq:pmf-beta-binomial} is usually unknown and will need to be
estimated from the survey data. An interpretation of \(\alpha\) is that
when it is low only a few plots are attractive to the observers and tend
to be selected, and when high, most plots are equally likely to be
chosen. VanderPlas et al define \(c\)-interesting plot to be if \(c\) or
more participants select the plot as the most different. The expected
number of plots selected at least \(c\) times, \(\mathrm{E}[Z_c]\), is
then calculated as

\begin{equation} \label{eq:c-interesting-expectation}
\mathrm{E}[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}

@@

Visual patterns of non-linearity are simulated using four different
orders of probabilist's Hermite polynomials (\(j = 2, 3, 6, 18\)). The
values of \(j\) were chosen so that distinct shapes of non-linearity
were included in the residual plot. These include ``U'', ``S'', ``M''
and ``triple-U'' shape as shown in Figure
\ref{fig:different-shape-of-herimite}. A greater value of \(j\) will
result in a curve with more turning points. It is expected that the
``U'' shape will be the easiest to detect, as the shape gets more
complex it will be harder to perceive in a scatterplot, particularly
when there is noise. Figure \ref{fig:different-sigma} shows the ``U''
shape for different amounts of noise (\(\sigma\)).

@@

Since \(\text{supp}(X) = [-1, 1]\), choosing \(a\) to be \(-1\), \(0\)
and \(1\) can generate ``left-triangle'', ``butterfly'' and
``right-triangle'' shapes as displayed in Figure
\ref{fig:different-shape-of-heter}. The term \((2 - |a|)\) maintains the
magnitude of residuals across different values of \(a\). Figure
\ref{fig:different-b} shows the butterfly shape as the ratio parameter
(\(b\)) is changed.

@@

\begin{itemize}
\tightlist
\item
  uniform, \(U(-1, 1)\),
\item
  normal, \(N(0, 0.3^2)\),
\item
  skewed, \(\text{lognormal}(0, 0.6^2)/3\), and
\item
  discrete, \(U\{1, 5\}\).
\end{itemize}

@@

\begin{equation}
E = \frac{1}{2}\left(\boldsymbol{\mu}_z'(\text{diag}(\boldsymbol{R}\sigma^2))^{-1}\boldsymbol{\mu}_z\right)
\label{eq:nonlin-effect-size}
\end{equation}

\noindent where \(\text{diag}(.)\) is the diagonal matrix constructed
from the diagonal elements of a matrix,
\(\boldsymbol{R} = \boldsymbol{I}_n - \boldsymbol{H}\) is the residual
operator,
\(\boldsymbol{H} = \boldsymbol{X}(\boldsymbol{X}'\boldsymbol{X})^{-1}\boldsymbol{X}'\)
is the hat matrix,
\(\boldsymbol{\mu}_z = \boldsymbol{R}\boldsymbol{Z}\boldsymbol{\beta}_z\)
is the expected values of residuals where \(\boldsymbol{Z}\) contain any
higher order terms of \(\boldsymbol{X}\) left out of the regression
equation, \(\boldsymbol{\beta}_z\) contains the corresponding
coefficients, and \(\sigma^2\boldsymbol{I}\) is the assumed covariance
matrix of the error term when \(H_0\) is true.

@@

Data collection used a total of 1152 lineups, and resulted in a total of
7974 evaluations from 443 participants. Roughly half corresponded to the
two models, non-linearity and heteroskedasticiy, and the three
collection periods had similar numbers of evaluations. Each participant
received two of the 24 attention check lineups which were used to filter
results of participants who were clearly not making an honest effort
(only 11 of 454). To estimate \(\alpha\) for calculating statistical
significance (see Section \ref{sig}) there were 720 evaluations of 36
null lineups. Neither the attention checks nor null lineups were used in
the subsequent analysis. The de-identified data, \texttt{vi\_survey}, is
made available in the R package, \texttt{visage}.

@@

In regression analysis, predictions are conditional on the observed
values of the predictors, that is, the conditional mean of the dependent
variable \(Y\) given the value of the independent variable \(X\),
\(\text{E}(Y|X)\). This is an often forgotten element of regression
analysis but it is important. Where \(X\) is observed, the distribution
of the \(X\) values in the sample, or consequently \(\hat{Y}\), may
affect the ability to read any patterns in the residual plots. The
effect of fitted value distribution on test performance is assess using
four different distributions of fitted values: uniform, normal, discrete
and lognormal (skewed). We expect that if \(\hat{Y}\) has a uniform
distribution, it is easier to read the relationship with the residuals.

@@

The R package \texttt{visage} is currently available at
\url{https://github.com/TengMCing/visage}.
